Identification of Direct Cherenkov Pixels using Boosted Decision Trees

### 2016-06-07 report/bdtarticle.tex
Backup 07/06/16

\begin{abstract}
When Cosmic Rays travel through the atmosphere, the primary particle will often emit some Direct Cherenkov light before the much brighter Extended Air Shower is generated. Unlike the broader air shower component of telescope images, all of the Direct Cherenkov light will usually be concentrated in a single pixel. A new method for identifying these Direct Cherenkov pixels was developed, relying on supervised machine learning. A set of Cosmic Ray telescope images was simulated, both with and without the Extended Air Shower background, for use in classifier training. With reference to the background-free telescope images, the Direct Cherenkov pixel in each corresponding full-shower telescope image could be reliably identified. A Boosted Decision Tree to identify these Direct Cherenkov pixels was trained, using data from the full-shower training images. The Boosted Decision Tree performance was then tested on a second set of full-shower telescope images, and compared to existing methods of Direct Cherenkov pixel identification. After applying cuts to the test sample, 37.3\% of all Direct Cherenkov pixels were accepted and correctly identified, representing a sevenfold increase in acceptance rates when compared to existing methods. The corresponding sample purity after cuts was 87.7\%, a slight increase on existing methods. When only high multiplicity Boosted Decision Tree events were considered, 11.8\% of all Direct Cherenkov pixels were identified, with a sample purity of 90.9\%, while the existing method was not able to find a single comparable high-multiplicity event. 
\end{abstract}
\section{Introduction}
Cosmic Rays will often emit Direct Cherenkov (DC) light in the upper atmosphere, before generating an Extended Air Shower (EAS) through interaction with the lower atmosphere. Analysis of the DC light from a shower can indicate primary particle energy, charge and position. It is consequently very useful for event reconstruction to measure the quantity and direction of DC light emission. There are numerous Telescope Arrays which currently image Cherenkov Light emission from Cosmic Rays in the atmosphere, such as the HESS, MAGIC and VERITAS experiments. In these telescope images, the DC light is usually concentrated in a single  \textquoteleft DC pixel'. Identifying this pixel is challenging, because the brighter EAS Cherenkov light background often overlaps with the DC pixel.

>>>>

\begin{abstract}
When Cosmic Rays travel through the atmosphere, the primary particle will often emit some Direct Cherenkov light before the much brighter Extended Air Shower is generated. Unlike the broader air shower component of telescope images, all of the Direct Cherenkov light will usually be concentrated in a single pixel. A new method for identifying these Direct Cherenkov pixels was developed, relying on supervised machine learning. A set of Cosmic Ray telescope images was simulated, both with and without the Extended Air Shower background, for use in classifier training. With reference to the background-free telescope images, the Direct Cherenkov pixel in each corresponding full-shower telescope image could be reliably identified. A Boosted Decision Tree to identify these Direct Cherenkov pixels was trained, using data from the full-shower training images. The Boosted Decision Tree performance was then tested on a second set of full-shower telescope images, and compared to existing methods of Direct Cherenkov pixel identification. After applying cuts to the test sample, 37.3\% of all Direct Cherenkov pixels were accepted and correctly identified, representing a sevenfold increase in acceptance rates when compared to existing methods. The corresponding sample purity after cuts was 87.7\%, a slight increase on existing methods. When only high multiplicity Boosted Decision Tree events were considered, 11.8\% of all Direct Cherenkov pixels were identified, with a sample purity of 90.9\%. The existing method was not able to find a single comparable high-multiplicity event. 
\end{abstract}
\section{Introduction}
Cosmic Rays will often emit Direct Cherenkov (DC) light in the upper atmosphere, before generating an Extended Air Shower (EAS) through interaction with the lower atmosphere. Analysis of the DC light from a shower can indicate primary particle energy, charge and position. It is consequently very useful for event reconstruction to measure the quantity and direction of DC light emission. There are numerous Telescope Arrays which currently image Cherenkov Light emission from Cosmic Rays in the atmosphere, such as the HESS, MAGIC and VERITAS experiments. In these telescope images, the DC light is usually concentrated in a single  \textquoteleft DC pixel'. Identifying this pixel is challenging, because the brighter EAS Cherenkov light background often overlaps with the DC pixel.
